A Methodology and Implementation of Automated Emissions Harmonization for Use in Integrated Assessment Models

### 2018-01-31 paper.tex
acs for ssps

\glspl{iam} are tools used to understand the complex
interactions between energy, economy, land use, water, and climate systems. \glspl{iam}
provide global projections of systemic change by dividing the world into a
number of representative regions (typically 10 to 30), the definition of which
is distinct for each model \cite{krey_global_2014}. Results from \glspl{iam} are
integral in a number of international studies, which notably include projections
of climate and economic futures. Recently, the \gls{iam} community has developed
scenarios based on the \glspl{ssp}
\cite{van_vuuren_energy_2017, fricko_marker_2017, fujimori_ssp3:_2017,
  calvin_ssp4:_2017, kriegler_fossil-fueled_2017} which quantify a variety of
potential global futures. The SSPs are designed to be used in research that
include \gls{esm} simulations, climate impact, adaptation and
climate mitigation studies \cite{vuuren_new_2013}.

While \glspl{iam} are implemented in myriad ways, including simulation and
optimization, the core inputs and outputs are similar across different
models. Modeling teams incorporate data on energy systems, land use, economics,
demographics and emissions sources and concentrations, among other data, in
order to provide consistent existing trajectories of modeled variables. The
models then provide estimates of future trajectories of these variables under
various socio-economic and technological assumptions as well as proposed policy
constraints, e.g., targets for future Greenhouse Gas (GHG) emissions.

The emissions trajectories calculated by \glspl{iam} are critical inputs for
ongoing, worldwide scientific community efforts in the \gls{cmip6}
\cite{eyring_overview_2016}, which is utilizing a number of marker SSP scenarios
developed by the \gls{iam} community (\gls{smip})
\cite{oneill_scenario_2016}. These trajectories are endogenously calculated by
modeling the individual technologies and sectors that contribute towards the
emissions of different air pollutants and GHGs as well as various mitigation
technologies. However, the historical emissions starting points of models can
differ by large amounts depending on the region, sector, and emissions species.

@@

The global climate change community has recently developed a new global
historical emissions data set for both anthropogenic emissions (i.e., the
\gls{ceds} \cite{hoesly_historical_2017} and
open-burning \gls{luluc} emissions
\cite{van_marle_historic_2017}) which, in conjunction with the SSP IAM
trajectories, will be used for climate-related modeling exercises of \gls{cmip6}.

@@

Climate modeling efforts have continued to progress, demanding increased spatial
and sectoral resolution from \glspl{iam}. Furthermore, a new generation of climate
scenarios which combines aspects of both the \glspl{rcp} and SSPs have been developed
in order to incorporate both physical and socio-economic detail. In order to
address the growing dimensionality of model outputs and support ongoing scenario
generation and analysis efforts while still providing a consistent and
scientifically rigorous harmonization procedure, an automated process for
determining harmonization methods is preferred. The use of an automated,
documented, and openly available harmonization mechanism additionally allows for
full procedural reproducibility and for direct participation by additional
modeling teams not involved in the original exercise.

@@

The value of $c_v$ is then tested against a threshold, $\tau_{c_v}$. To
determine this threshold, an analysis of the recent \gls{ceds} and \gls{luc} historical data
has been performed. Figure \ref{fig:cov} shows the distribution of \gls{luc} $c_v$s
and non-\gls{luc} $c_v$s as determined for historical data aggregated to the model
regions of 5 different \glspl{iam} involved in the SSP process: AIM-CGE, IMAGE, GCAM4,
MESSAGE-GLOBIOM, and REMIND-MAGPIE \TODO{add refs}, each of which have varying
definitions of native model regions comprising different collections of
countries. Therefore, each data point comprising Figure \ref{fig:cov} represents
a realization of $c_v$ for a single combination of native model region, sector,
and emissions species \footnote{A full listing of all sectors and species is
  presented in the case-study discussion in Section \ref{sec:results}, Table
  \ref{tab:sp}}. A threshold value of $\tau_{c_v} = 20$ has been chosen based on
these observations as it optimally divides the two distributions. Importantly,
tails of the \gls{luc} and non-\gls{luc} overlap, thus there are both false positives (~7\%
of non-\gls{luc} trajectories) and false negatives (~10\% of \gls{luc}
trajectories). However, as any regional definition is model dependent and thus
any regional aggregation is possible an automated detection mechanism is
necessary.

@@

Finally, consideration is taken with respect to the relative difference between
the historic and model values in the harmonization time period. In order to
investigate the possible values that these relative differences can take, the
\gls{iam} values used in the SSP and (ongoing) \gls{cmip6} inter-comparison
exercises are used. A distribution of these differences for all models in the
study is presented in Figure \ref{fig:dh}. Given the available data, a threshold
value of $\tau_{dH} = 50$\% was chosen to be used as a default in \code{aneris}.

@@

In order to show a representative cross section of the performance of the
\code{aneris} harmonization procedure, we focus on the harmonization of results
of the \gls{iam} MESSAGE-GLOBIOM \cite{fricko_marker_2017}. Two scenarios from the SSP
scenario library \cite{Riahi2017153,Rao2017346} are presented. The SSP2, or
``middle of the road'', scenario (referred to as SSP2-Ref) is chosen to be
analyzed because MESSAGE-GLOBIOM is the marker scenario for this SSP. This SSP2
scenario lies between two \glspl{rcp}, 6 and 8.5, with a radiative forcing level of
approximately 6.5 Wm$^{-2}$. We additionally present the results for the
SSP2-based mitigation scenario leading to a radiative forcing of 4.5 Wm$^{-2}$
(referred to as SSP2-4.5). The SSP2-45 scenario is chosen because mitigation
technologies and policies are enacted causing a general reduction in pollutants
and GHGs, including (eventual) negative \cotwo emissions in some regions and
sectors due to carbon capture and sequestration and afforestation. A scenario in
which negative emissions play a role in mitigation strategies is particularly
important because of the sensitivity of key indicators, such as end-of-century
radiative forcing (which is used to estimate mean global temperature increase),
to the timing and magnitude of net-zero and total negative \cotwo
emissions. Therefore, these two scenarios represent two relative extreme cases
in the use of a harmonization approach and thus provide a case study as to its
general applicability. Figure \ref{fig:kyoto} shows the different trends of
Kyoto Gases in each scenarios.

@@

MESSAGE-GLOBIOM includes a representation of 11 distinct regions which can be
mapped directly to the 5-region definition used in the \glspl{rcp}. Historical data is
taken from previously described \gls{luc} and anthropogenic sources, which comprise 10
separate pollutant and GHG species and 12 sectors shown in Table \ref{tab:sp}. A
total of 970 distinct trajectories were harmonized for each scenario, and
therefore 1940 trajectories were harmonized in total. \noxx generated from the
Energy sector provides an example of an emissions species and sector in which
all regions were satisfactorily harmonized with the default methods. Figure
\ref{fig:nox} shows the results of harmonization in Asia, and Table
\ref{tab:nox} describes the parameters that underlie the choice of method for
each harmonized trajectory.

>>>>

\glspl{iam} are tools used to understand the complex interactions between
energy, economy, land use, water, and climate systems. \glspl{iam} provide
global projections of systemic change by dividing the world into a number of
representative regions (typically 10 to 30), the definition of which is distinct
for each model \cite{krey_global_2014}. Results from \glspl{iam} are integral in
a number of international studies, which notably include projections of climate
and economic futures. Recently, the \gls{iam} community has developed scenarios
based on the \glspl{ssp} \cite{van_vuuren_energy_2017, fricko_marker_2017,
  fujimori_ssp3:_2017, calvin_ssp4:_2017, kriegler_fossil-fueled_2017} which
quantify a variety of potential global futures. The \glspl{ssp} are designed to
be used in research that include \gls{esm} simulations, climate impact,
adaptation and climate mitigation studies \cite{vuuren_new_2013}.

While \glspl{iam} are implemented in myriad ways, including simulation and
optimization, the core inputs and outputs are similar across different
models. Modeling teams incorporate data on energy systems, land use, economics,
demographics and emissions sources and concentrations, among other data, in
order to provide consistent existing trajectories of modeled variables. The
models then provide estimates of future trajectories of these variables under
various socio-economic and technological assumptions as well as proposed policy
constraints, e.g., targets for future \gls{ghg} emissions.

The emissions trajectories calculated by \glspl{iam} are critical inputs for
ongoing, worldwide scientific community efforts in the \gls{cmip6}
\cite{eyring_overview_2016}, which is utilizing a number of marker \gls{ssp} scenarios
developed by the \gls{iam} community (\gls{smip})
\cite{oneill_scenario_2016}. These trajectories are endogenously calculated by
modeling the individual technologies and sectors that contribute towards the
emissions of different air pollutants and \glspl{ghg} as well as various
mitigation technologies. However, the historical emissions starting points of
models can differ by large amounts depending on the region, sector, and
emissions species.

@@

The global climate change community has recently developed a new global
historical emissions data set for both anthropogenic emissions (i.e., the
\gls{ceds} \cite{hoesly_historical_2017} and open-burning \gls{luluc} emissions
\cite{van_marle_historic_2017}) which, in conjunction with the \gls{ssp}
\gls{iam} trajectories, will be used for climate-related modeling exercises of
\gls{cmip6}.

@@

Climate modeling efforts have continued to progress, demanding increased spatial
and sectoral resolution from \glspl{iam}. Furthermore, a new generation of climate
scenarios which combines aspects of both the \glspl{rcp} and \glspl{ssp} have been developed
in order to incorporate both physical and socio-economic detail. In order to
address the growing dimensionality of model outputs and support ongoing scenario
generation and analysis efforts while still providing a consistent and
scientifically rigorous harmonization procedure, an automated process for
determining harmonization methods is preferred. The use of an automated,
documented, and openly available harmonization mechanism additionally allows for
full procedural reproducibility and for direct participation by additional
modeling teams not involved in the original exercise.

@@

The value of $c_v$ is then tested against a threshold, $\tau_{c_v}$. To
determine this threshold, an analysis of the recent \gls{ceds} and \gls{luc} historical data
has been performed. Figure \ref{fig:cov} shows the distribution of \gls{luc} $c_v$s
and non-\gls{luc} $c_v$s as determined for historical data aggregated to the model
regions of 5 different \glspl{iam} involved in the \gls{ssp} process: AIM-CGE, IMAGE, GCAM4,
MESSAGE-GLOBIOM, and REMIND-MAGPIE \TODO{add refs}, each of which have varying
definitions of native model regions comprising different collections of
countries. Therefore, each data point comprising Figure \ref{fig:cov} represents
a realization of $c_v$ for a single combination of native model region, sector,
and emissions species \footnote{A full listing of all sectors and species is
  presented in the case-study discussion in Section \ref{sec:results}, Table
  \ref{tab:sp}}. A threshold value of $\tau_{c_v} = 20$ has been chosen based on
these observations as it optimally divides the two distributions. Importantly,
tails of the \gls{luc} and non-\gls{luc} overlap, thus there are both false positives (~7\%
of non-\gls{luc} trajectories) and false negatives (~10\% of \gls{luc}
trajectories). However, as any regional definition is model dependent and thus
any regional aggregation is possible an automated detection mechanism is
necessary.

@@

Finally, consideration is taken with respect to the relative difference between
the historic and model values in the harmonization time period. In order to
investigate the possible values that these relative differences can take, the
\gls{iam} values used in the \gls{ssp} and (ongoing) \gls{cmip6} inter-comparison
exercises are used. A distribution of these differences for all models in the
study is presented in Figure \ref{fig:dh}. Given the available data, a threshold
value of $\tau_{dH} = 50$\% was chosen to be used as a default in \code{aneris}.

@@

In order to show a representative cross section of the performance of the
\code{aneris} harmonization procedure, we focus on the harmonization of results
of the \gls{iam} MESSAGE-GLOBIOM \cite{fricko_marker_2017}. Two scenarios from the \gls{ssp}
scenario library \cite{Riahi2017153,Rao2017346} are presented. The SSP2, or
``middle of the road'', scenario (referred to as SSP2-Ref) is chosen to be
analyzed because MESSAGE-GLOBIOM is the marker scenario for this \gls{ssp}. This SSP2
scenario lies between two \glspl{rcp}, 6 and 8.5, with a radiative forcing level of
approximately 6.5 Wm$^{-2}$. We additionally present the results for the
SSP2-based mitigation scenario leading to a radiative forcing of 4.5 Wm$^{-2}$
(referred to as SSP2-4.5). The SSP2-45 scenario is chosen because mitigation
technologies and policies are enacted causing a general reduction in pollutants
and \glspl{ghg}, including (eventual) negative \cotwo emissions in some regions and
sectors due to carbon capture and sequestration and afforestation. A scenario in
which negative emissions play a role in mitigation strategies is particularly
important because of the sensitivity of key indicators, such as end-of-century
radiative forcing (which is used to estimate mean global temperature increase),
to the timing and magnitude of net-zero and total negative \cotwo
emissions. Therefore, these two scenarios represent two relative extreme cases
in the use of a harmonization approach and thus provide a case study as to its
general applicability. Figure \ref{fig:kyoto} shows the different trends of
Kyoto Gases in each scenarios.

@@

MESSAGE-GLOBIOM includes a representation of 11 distinct regions which can be
mapped directly to the 5-region definition used in the \glspl{rcp}. Historical data is
taken from previously described \gls{luc} and anthropogenic sources, which comprise 10
separate pollutant and \gls{ghg} species and 12 sectors shown in Table \ref{tab:sp}. A
total of 970 distinct trajectories were harmonized for each scenario, and
therefore 1940 trajectories were harmonized in total. \noxx generated from the
Energy sector provides an example of an emissions species and sector in which
all regions were satisfactorily harmonized with the default methods. Figure
\ref{fig:nox} shows the results of harmonization in Asia, and Table
\ref{tab:nox} describes the parameters that underlie the choice of method for
each harmonized trajectory.
